---
title: "Beach Volleyball(2020_21) - TidyTuesday"
author: "[Allie Burns](https://allie-burns.github.io/)"
date: "`r Sys.Date()`"

output:
  html_document:
    toc: false
    code_folding: hide
editor_options:
  chunk_output_type: console
---


```{r setup, include=FALSE}
## Set options for knitr. Show code chunks but not warnings.
knitr::opts_chunk$set(echo = TRUE, warning = F)
```

## Loading and exploring the data
This week's data is an exhaustive list of results from volleyball tournaments since the year 2000. There are so many questions that can be answered with this data. I could look at player statistics, tournament statistics or statistics within each round

```{r prep, message=FALSE, warning=FALSE}
##
library(skimr)
library(tidyverse)
## Load in the libraries
devtools::install_github("thebioengineer/tidytuesdayR")

## Install tidy data
tuesdata <- tidytuesdayR::tt_load(2020, week = 21)
vb <- tuesdata$vb_matches

## check the data
skim(vb)

```

First, I want to see if games become more aggressive with each round.  For now, I'll just focus on the women.

```{r save, include=FALSE}

tt <- vb %>%
    filter(gender == 'W') %>%
    group_split(tournament) 


pdf(file = "../../plots/2020_20/20200512_VolcanoEruptions.pdf", width = 12, height = 7)
world_volcanos
all_population
highest_population
dev.off()

png(filename = "../../plots/2020_20/20200512_VolcanoEruptions.png", width = 900, height = 500)
world_volcanos
dev.off()
```


## Today I learned:
+ __skimr__ : This is such a cool tool for data exploration and I'm suprised that I had never heard of it before.  It was an comprehensive method to get an idea of what data was available without having to run summary statistics for every single column.  And after looking at all of the summary statistics, I had many more questions that I thought the data could answer


## Session Info
```{r session-info}
 sessionInfo()
```

